---
title: "Upset plot and heatmap example for RNAseq data"
output: html_notebook
author: 'Christina Fitzsimmons'
date: 'Last updated 2021-07-04'
---
# Introduction
This notebook shows an example of and upset plot and a heatmap--two common types of plots that show the similarities and differences between large data sets. The example below is from a different project of mine, but the concept of converting RNAseq data into plots is similar. 

```{r Libraries}
library(DESeq2)
library(biomaRt)
library(tidyverse)
library(pheatmap)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(gplots)
library(EnhancedVolcano)
library(UpSetR)
library(plyr)
library(reshape2)
library(dplyr)
library(plotly)
```
# 1. Generating the upset plot
# 1.1 Normoxia Data Only
Checking to determine if we get same number of genes in upset plot when only normoxia data is loaded as compared to all 18 data sets. Might be different!

```{r Generating the Upset Plot}
# take a dataframe from the DEseq2 output
# Filter dataframes for padj only (res_sig)
# select the geneID column 
# make a new variable, set. All geneIDs on this list get a value of 1 (binary 1/0 needed for upset plot)
MutvPar_padjF <- as.data.frame(res_MutvParent_sig) %>%
  rownames_to_column(var = "geneID")%>% 
  dplyr::select(geneID) %>%
  dplyr::mutate(set="MutvPar", val=1)

WTvPar_padjF <- as.data.frame(res_WTvParent_sig) %>%
  rownames_to_column(var = "geneID") %>%
  dplyr::select(geneID) %>%
  dplyr::mutate(set="WTvPar", val=1)

MutvWT_padj <- as.data.frame(res_MutvWT_sig) %>%
  rownames_to_column(var = "geneID") %>%
  dplyr::select(geneID) %>%
  dplyr::mutate(set="MutvWT", val=1)

padj_long <- bind_rows(MutvPar_padjF, WTvPar_padjF, MutvWT_padj) #generate 1 large dataframe with values from all 3 comparisons

# Making the Upset Plot
# transition the dataframe to a wide format
test1<-acast(padj_long,geneID~set,value.var="val",fill=0)

# force as dataframe
test1<-as.data.frame(test1)

# 1st column must be a name
test1$name=rownames(test1)

test2 <- test1 %>%
  dplyr::select(name, MutvPar, WTvPar, MutvWT)

# Generate the upset plot and print out to pdf. 
upset(test2,nsets = 3, nintersects = NA, order.by = "freq")
pdf('2021.01.24_UpsetPlot_Normoxia-only.pdf') # name of pdf to save
upset(test2,nsets = 6, nintersects = NA, order.by = "freq")
dev.off()

# exclude or comment out lines 67 and 69 if you don't wish to save a pdf of the plot. 

```
## 1.2 Extracting a list of genes to make a heatmap. 
```{r Generating the filtered lists for the heatmap}

# Doing this the slower way with dplyr anti-joins. 
WTunique<- anti_join(WTvPar_padjF, MutvWT_padj, by=c("geneID")) # Keep genes only upregulated in WT vs Parent, not Mut vs WT (controls for cancer genes)
WTunique2 <- anti_join(WTunique, MutvPar_padjF, by = c("geneID")) %>% dplyr::select(geneID) # Keep those genes only upregulated in WT, not Mut vs Parent (controls for lentiviral expression)

head(WTunique2)

WTvPar_df <- as.data.frame(res_WTvParent_sig) %>% rownames_to_column(var = "geneID") %>% dplyr::select(geneID, log2FoldChange, padj)
MutvPar_df <- as.data.frame(res_MutvParent) %>% rownames_to_column(var = "geneID") %>% dplyr::select(geneID, log2FoldChange, padj)
MutvWT_df <- as.data.frame(res_MutvWT) %>% rownames_to_column(var = "geneID") %>% dplyr::select(geneID, log2FoldChange, padj)

# make a dataframe of the genes unique to WT vs Parent
WTvPar_717 <- dplyr::semi_join(WTvPar_df, WTunique2) %>%
  dplyr::rename(WTvPar_padj=padj, WTvPar_l2fc=log2FoldChange)

MutvPar_717 <- dplyr::semi_join(MutvPar_df, WTunique2) %>%
  dplyr::rename(MutvPar_padj=padj, MutvPar_l2fc=log2FoldChange)

MutvWT_717 <- dplyr::semi_join(MutvWT_df, WTunique2) %>% 
  dplyr::rename(MutvWT_padj=padj, MutvWT_l2fc=log2FoldChange)

# print that dataframe of WT vs Par
d1 <- merge(WTvPar_717, MutvWT_717)
d2 <- merge(d1, MutvPar_717)
#write.csv(d2, "./2021.01.25_UpsetPlot_717genes-unique-to_WTvPar_padj-l2fc.csv")

# Many of the genes here have a padj of 0.99 or very close to one
# practically either this means that they are not worth comparing--because they are the same
# OR, an alternative explanation, that we have low number of counts. 

# semi_join to create the heatmap list and then write out
df3 <- semi_join(MutvWT_padj, WTvPar_padjF, by = "geneID")
heatmap_gene2 <- anti_join(df3, MutvPar_padjF, by= "geneID") 
#write.csv(heatmap_gene2, "2021.01_upsetR_heatmapgenes_1048.csv")

```
# 2 Heatmap of Normoxia 262
Now going to try making the heatmap with the above list. Might need to make a cutoff, we'll see if 1048 is too many genes to plot...
## 2.1 Heatmap of upset plot genes
```{r Heatmap plotting}

# This is a useful reference!
# https://jokergoo.github.io/ComplexHeatmap-reference/book/upset-plot.html#upset-plots-as-heatmaps

# get the genes
sigGenes <- heatmap_gene2 %>% 
  pull("geneID")


# filter the data for the genes on our list on line 121
plotDat <- vst(dds)[sigGenes,] %>% 
  assay()
z.mat <- t(scale(t(plotDat), center=TRUE, scale=TRUE))
myPalette <- c("#d7191c", "#ffffbf", "#2c7bb6") #defines colors used for plotting
myRamp = colorRamp2(c(-2, 0, 2), myPalette) # use my colors and plot based on z-score

Heatmap(z.mat, name = "z-score",
        col = myRamp,            
        show_row_name = FALSE,
        cluster_columns = TRUE)

# write out a pdf of the heatmap
#pdf('2021.01.24_Heatmap_Normoxia-only_Par-Mut_WT-1048genes.pdf')
#Heatmap(z.mat, name = "z-score",
#        col = myRamp,            
#        show_row_name = FALSE,
#        cluster_columns = TRUE)
#dev.off()
```
## 2.2 Heatmap of ONLY Mut vs WT genes
```{r Heatmap of Mut v WT only}

# get the genes from MutvWT with padj < 0.05
# Based on the upset plot, this should be 1284
sigGenes2 <- MutvWT_padj %>% 
  pull("geneID")


# filter the data for the top 200 by padj in the LRT test
plotDat <- vst(dds)[sigGenes2,] %>% 
  assay()
z.mat <- t(scale(t(plotDat), center=TRUE, scale=TRUE))
# colour palette
myPalette <- c("red3", "ivory", "blue")
myRamp = colorRamp2(c(-2, 0, 2), myPalette)

Heatmap(z.mat, name = "z-score",
        col = myRamp,            
        show_row_name = FALSE,
        cluster_columns = TRUE)

# write out a pdf of the heatmap
#pdf('2021.01.25_Heatmap_Normoxia-only_MutvWT_padj05_1284genes.pdf')
#Heatmap(z.mat, name = "z-score",
#        col = myRamp,            
#        show_row_name = FALSE,
#        cluster_columns = TRUE)
#dev.off()

# brewer.pal(n=3, name = 'RdYlBu')

```
# 3. Volcano Plot
Generating volcano plots of the MutvWT and MutvPar for our supplemental data. 
```{r Volcano Plot}

volcano <- as.data.frame(res_noNA_MutvWT) %>%
  rownames_to_column(var="geneID") %>%
  dplyr::mutate(logP = -1*log10(padj)) 

test_volcano <- ggplot(volcano, aes(x=log2FoldChange, y=logP)) +
  geom_point()+
  theme_bw() +
  xlim(-5,5) + # set x-limits on plot so both appear on the same scale (no genes cut off here)
  geom_vline(xintercept = -1, linetype="dashed") + geom_vline(xintercept = 1, linetype = "dashed") + # vertical dashed lines at significant log2foldchange
  geom_hline(yintercept = 1.30, linetype = "dashed") + # horizontal dashed line at significant -10log(padj) values
  labs(x = "Log2FoldChange (Mut/WT)", y = "-Log10(padj)", title = "Volcano plot of UOK262 Normoxia RNAseq") 

test_volcano # plots the volcano object. 

#pdf('2021.01.25_VolcanoPlot_MutvWT_Normoxia_RNAseq.pdf')
#test_volcano
#dev.off()


my_fav_gene <- as.data.frame(res_MutvParent) %>%
  rownames_to_column(var="geneID") %>%
  dplyr::filter(geneID == "ENSG00000091483.6") %>%
  dplyr::mutate(padj = 5.61E-301) %>%
  dplyr::mutate (logP = -1*log10(padj))

volcano2 <- as.data.frame(res_noNA_MutvParent) %>%
  rownames_to_column(var="geneID") %>%
  dplyr::mutate(logP = -1*log10(padj)) 

test_volcano2 <- ggplot(volcano2, aes(x=log2FoldChange, y=logP, label=geneID)) +
  geom_point()+
  geom_point(data=my_fav_gene, color="red") + #specifically adding FH to the plot manually. Shows that this gene is really only one sig. upregulated. 
  theme_bw() +
  xlim(-5,5) +
  geom_vline(xintercept = -1, linetype="dashed") + geom_vline(xintercept = 1, linetype = "dashed") +
  geom_hline(yintercept = 1.30, linetype = "dashed") +
  labs(x = "Log2FoldChange (Mut/Par)", y = "-Log10(padj)", title = "Volcano plot of UOK262 Normoxia RNAseq") 

test_volcano2

# print the volcano object to a pdf. 
pdf('2021.01.31_VolcanoPlot_MutvPar_Normoxia_RNAseq.pdf')
test_volcano2
dev.off()