---
title: "Cowsay Practice Notebook"
output: html_notebook
author: Ryan Gensler
date: 2021-06-09
---
# Introduction
This notebook serves as an introduction to Rnotebooks using the fun and silly `cowsay` library which makes easy to print messages or warnings with various animals. You'll notice we have 2 kinds of blocks in this notebook. Whitespace, where we can type longer notes and comments to ourselves much like a digial notebook, and code blocks (the grey blocks below)
```{r}
# The type of codeblock (R, bash, python, etc) is indicated in the upper left. 
# To run the code in the codeblock, click the green arrow in the upper right. 
# This codeblock loads the library.
library(cowsay)
# Anything I write after the pound symbol is ignored as a comment. 
```
# Animals
There are over 40 animals that you can access in `cowsay`. You can list the available animals with the following command:
```{r}
sort(names(animals))
```
In `cowsay` you can call on any animal in the package and make it say whatever you want. For example:
```{r}
say("why did the chicken cross the road", "chicken")
```
```{r}
say("boo!", "ghost")
```
# Special commands
## catfact
`catfact` will print a new fact about cats each time the code is run. Try running this block a few times and reading the new facts
```{r}
say("catfact", "cat")
```
## time
`time` is another special command that will print the current date and time 
```{r}
library(cowplot)
say('time')
```
## signbunny
`signbunny` is a commonly used internet meme 
```{r}
say(by='signbunny')
```
# Color
You can also add color to the animals by using the crayon package. All colors() are all supported as are any rgb values that evaluate to a string.
```{r}
library(jsonlite)
say(what = "fortune",
    by = "squirrel",
    what_color = "#FF0099",
    by_color = "red")
```
#Practice Problems:
1) Change the author name and the date at the top of the notebook and save the notebook with a new name
2) Have a chicken tell a road crossing joke
3) Have a green yoda say something about the dark side
4) Have a blue shark say the date of the next upcoming shark week
5) Have your favorite animal (from the list) say your name

 ----- 
Bad is the darkside 
 ------ 
    \   
     \
                   ____
                _.' :  `._
            .-.'`.  ;   .'`.-.
   __      / : ___\ ;  /___ ; \      __
  ,'_ ""--.:__;".-.";: :".-.":__;.--"" _`,
  :' `.t""--.. '<@.`;_  ',@>` ..--""j.' `;
       `:-.._J '-.-'L__ `-- ' L_..-;'
          "-.__ ;  .-"  "-.  : __.-"
             L ' /.------.\ ' J
             "-.   "--"   .-"
             __.l"-:_JL_;-";.__
         .-j/'.;  ;""""  / .'\"-.
         .' /:`. "-.:     .-" .';  `.
      .-"  / ;  "-. "-..-" .-"  :    "-.
  .+"-.  : :      "-.__.-"      ;-._   \
  ; \  `.; ;                    : : "+. ;
  :  ;   ; ;                    : ;  : \:
  ;  :   ; :                    ;:   ;  :
  : \  ;  :  ;                  : ;  /  ::
  ;  ; :   ; :                  ;   :   ;:
  :  :  ;  :  ;                : :  ;  : ;
  ;\    :   ; :                ; ;     ; ;
  : `."-;   :  ;              :  ;    /  ;
 ;    -:   ; :              ;  : .-"   :
  :\     \  :  ;            : \.-"      :
  ;`.    \  ; :            ;.'_..--  / ;
  :  "-.  "-:  ;          :/."      .'  :
   \         \ :          ;/  __        :
    \       .-`.\        /t-""  ":-+.   :
     `.  .-"    `l    __/ /`. :  ; ; \  ;
       \   .-" .-"-.-"  .' .'j \  /   ;/
        \ / .-"   /.     .'.' ;_:'    ;
  :-""-.`./-.'     /    `.___.'
               \ `t  ._  /  bug
                "-.t-._:'
  
> say("To get to the other side", chicken)
Error in match.arg(by, c(choices = names(animals), "rms", "random")) : 
  object 'chicken' not found
> say("To get to the other side", "chicken")

 ----- 
To get to the other side 
 ------ 
    \   
     \
         _
       _/ }
      `>' \
      `|   \
       |   /'-.     .-.
        \'     ';`--' .'
         \'.    `'-./
          '.`-..-;`
            `;-..'
            _| _|
            /` /` [nosig]
  
> say("The next shark week is July 11-18th", "shark")

 -------------- 
The next shark week is July 11-18th 
 --------------
    \
      \
        \
              /""-._
              .       '-,
               :          '',
                ;      *     '.
                 ' *         () '.
                   \               \
                    \      _.---.._ '.
                    :  .' _.--''-''  \ ,'
        .._           '/.'             . ;
        ; `-.          ,                \'
         ;   `,         ;              ._\
          ;    \     _,-'                ''--._
          :    \_,-'                          '-._
          \ ,-'                       .          '-._
          .'         __.-'';            \...,__       '.
        .'      _,-'        \              \   ''--.,__  '\
        /    _,--' ;         \              ;           \^.}
        ;_,-' )     \  )\      )            ;
             /       \/  \_.,-'             ;
            /                              ;
         ,-'  _,-'''-.    ,-.,            ;      PFA
      ,-' _.-'        \  /    |/'-._...--'
     :--``             )/
  '
  
> say("hello Ryan", by = "monkey", type = "warning")
Warning message:
In say("hello Ryan", by = "monkey", type = "warning") : 
 ------------- 
hello Ryan 
 -------------- 
              \   
               \  
                \

                  .="=.
                _/.-.-.\_     _
               ( ( o o ) )    ))
                |/  "  \|    //
                 \'---'/    //
           jgs   /`"""`\\  ((
                / /_,_\ \\  \\
                \_\_'__/  \  ))
                /`  /`~\   |//
               /   /    \  /
          ,--`,--'\/\    /
          '-- "--'  '--'
Done! 