---
title: "ColinCowsayPractice"
output: html_notebook
author: Colin Fischer
date: 2021-06-09
---
# Introduction
This notebook serves as an introduction to Rnotebooks using the fun and silly `cowsay` library which makes easy to print messages or warnings with various animals. You'll notice we have 2 kinds of blocks in this notebook. Whitespace, where we can type longer notes and comments to ourselves much like a digial notebook, and code blocks (the grey blocks below)
```{r}
# The type of codeblock (R, bash, python, etc) is indicated in the upper left. 
# To run the code in the codeblock, click the green arrow in the upper right. 
# This codeblock loads the library.
library(cowsay)
# Anything I write after the pound symbol is ignored as a comment. 
```
# Animals
There are over 40 animals that you can access in `cowsay`. You can list the available animals with the following command:
```{r}
sort(names(animals))
```
In `cowsay` you can call on any animal in the package and make it say whatever you want. For example:
```{r}
say("why did the chicken cross the road", "chicken")
```
```{r}
say("boo!", "ghost")
```
# Special commands
## catfact
`catfact` will print a new fact about cats each time the code is run. Try running this block a few times and reading the new facts
```{r}
say("catfact", "cat")
```
## time
`time` is another special command that will print the current date and time 
```{r}
library(cowplot)
say('time')
```
## signbunny
`signbunny` is a commonly used internet meme 
```{r}
say(by='signbunny')
```
# Color
You can also add color to the animals by using the crayon package. All colors() are all supported as are any rgb values that evaluate to a string.
```{r}
library(jsonlite)
say(what = "fortune",
    by = "squirrel",
    what_color = "#FF0099",
    by_color = "red")
```
#Practice Problems:
1) Change the author name and the date at the top of the notebook and save the notebook with a new name
2) Have a chicken tell a road crossing joke
3) Have a green yoda say something about the dark side
4) Have a blue shark say the date of the next upcoming shark week
5) Have your favorite animal (from the list) say your name
```{r}
```


```{r}
say("why did the chicken cross the road", "chicken")
```
```{r}
say("to escape the KFC", "chicken")
```
```{r}
say(what = "mysterious as the dark side of the moon Mulan is",
    by = "yoda",
    by_color = "green")
```
```{r}
say(what = "I don't have a tv, computer or phone, how am I supposed to know!",
    by = "shark",
    by_color = "blue")
```
```{r}
say(what = "Colin",
    by = "cow",
    by_color = "purple")
```


Done! 